How to publish in Biological Sciences: a guide for the uninitiated

### 2022-09-08 2.04-whichjournal.Rmd
# new chapter on APCs

### Indexing

You want other people to be able to find the work that you publish, and so selecting a journal that is already indexed in one of the major literature databases (Web of Science or Scopus) is important. Be aware that when new journals are indexed, they are usually done with their entire back catalogue. Thus, if a journal advertises that it will be listed by one of major literature databases, your submission will likely be listed eventually, even if not immediately.

@@

Journals have different hierarchies of scope. Some journals attempt to take on the full gambit of science (e.g. *Science* and *Nature*) while others are only interested in a particular taxonomic group. In general, the impact factor of the journal is likely to be linked to the diversity of the scope. This is not always the case. There are some taxonomically specific journals with high impact factors, and there are some general journals with low impact factors.

@@

In an ideal world, there wouldn’t be any more barriers to you publishing your contribution to the collective of scientific knowledge. Not only is it not an ideal world, but I would argue that there has never been a less ideal time for publishing science. Greedy publishers have taken publicly funded science and made the content creators pay for it time and again. Read more on this [later in the book](#paywall).

>>>>

### Indexing {#indexing}

You want other people to be able to find the work that you publish, and so selecting a journal that is already indexed in one of the major literature databases (Web of Science or Scopus) is important. You can read more about literature databases and what you can do with them [here](https://howtowriteaphd.org/databases.html). You can search one of these databases to determine whether or not your journal in indexed there, but many of the journals advertise on their home pages where they are indexed. Be aware that when new journals are indexed, they are usually done with their entire back catalogue. Thus, if a journal advertises that it will be listed by one of major literature databases, your submission will likely be listed eventually, even if not immediately.

@@

Journals have different hierarchies of scope. Some journals attempt to take on the full gambit of science (e.g. *Science* and *Nature*) while others are only interested in a particular taxonomic group. In general, the impact factor of the journal is likely to be linked to the diversity of the scope. This is not always the case. There are some taxonomically specific journals with high impact, and there are some general journals with low impact.

[JANE (Journal/Author Name Estimator)](https://jane.biosemantics.org/) is a useful tool to help you determine an appropriate target journal. You can enter your title, abstract or keywords into a search box and get a list of journal names with articles having similar titles, together with links to their [PubMed](https://pubmed.ncbi.nlm.nih.gov/) records. Because JANE is only linked to PubMed, it may be of limited use in your area of Biological Sciences. However, you can use similar search methods in other literature databases (Web of Sciences, Dimensions, Scopus, etc.) to get similar results.

@@

In an ideal world, there wouldn’t be any more barriers to you publishing your contribution to the collective of scientific knowledge. Not only is it not an ideal world, but I would argue that there has never been a less ideal time for publishing science. Read more on this [later in the book](#paywall).
